Analysis of Shootings and Firearm Discharges in Toronto

### 2024-11-29 paper/paper.qmd
update paper

@@

-   **Weighted Score**: Combines the number of deaths and injuries, weighting deaths twice as heavily(weighted_score = death \* 2 + injuries).

@@

### Variable Importance
The summary function highlighted the relative importance of each predictor shown in @fig-importance. Notably, the neighborhood variable (neighbourhood_158) dominated the model, indicating significant spatial variation in shooting severities across Toronto. The occ_dow and division showed moderate influence, suggesting that day of the week and police division also contribute to variations in incident severity, albeit to a lesser extent.
```{r}
#| warning: false
#| message: false
#| label: fig-importance
#| fig_caption: true
#| fig-cap: "Bar chart shows the importance of the predictor variables"
#| echo: false

@@

# Plot variable importance
summary(model, n.trees = 500, plotit = TRUE)
```
### Model Performance
The model’s accuracy was assessed using the Root Mean Square Error (RMSE), calculated at 2.173830. This metric quantifies the average magnitude of the model’s prediction errors, providing a measure of predictive accuracy.

### Result Interpretation
The results underline the critical influence of geographic location (neighbourhood_158) on the severity of shooting incidents, implying that certain areas are systematically associated with higher severity scores. Temporal factors like the day of the week (occ_dow) and specific police divisions also play roles but are less pronounced compared to spatial factors.

# Results

The model highlighted significant predictors and their relative influence on the severity of shootings. `Division` emerged as a particularly influential predictor, indicating that certain areas within Toronto experience distinctly different levels of shooting severity.

@@

Based on the model's findings, policy recommendations are proposed to enhance public safety and optimize police resource allocation.

@@

This study provides a comprehensive analysis of the factors affecting the severity of shootings in Toronto, offering actionable insights for policymakers and law enforcement agencies.

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(gbm)
library(caret)
library(ggplot2)

#### Load saved model ####
model <- readRDS("models/first_model.rds")

# Load the cleaned and preprocessed data for further analysis or prediction
analysis_data <- read_csv("../data/02-analysis_data/analysis_data.csv")

@@

-   `Weighted Score`: Combines the number of deaths and injuries, weighting deaths twice as heavily(weighted_score = death \* 2 + injuries).

@@

### Variable Importance

The summary function highlighted the relative importance of each predictor shown in @fig-importance. Notably, the neighborhood variable (neighbourhood_158) dominated the model, indicating significant spatial variation in shooting severities across Toronto. The occ_dow and division showed moderate influence, suggesting that day of the week and police division also contribute to variations in incident severity, albeit to a lesser extent.

```{r}
#| warning: false
#| message: false
#| label: fig-importance
#| fig_caption: true
#| fig-cap: "Bar chart shows the importance of the predictor variables"
#| echo: false

@@

# Plot variable importance
summary(model, n.trees = 500, plotit = TRUE)
```

### Model Performance

The model’s accuracy was assessed using the Root Mean Square Error (RMSE), calculated at 2.173830. This metric quantifies the average magnitude of the model’s prediction errors, providing a measure of predictive accuracy.

### Result Interpretation

The results underline the critical influence of geographic location (neighbourhood_158) on the severity of shooting incidents, implying that certain areas are systematically associated with higher severity scores. Temporal factors like the day of the week (occ_dow) and specific police divisions also play roles but are less pronounced compared to spatial factors.

@@

Based on the model’s findings, it is recommended that public safety measures and police resources be strategically focused on specific neighborhoods and times identified as high-risk, enhancing targeted interventions and preventive strategies. These efforts should prioritize the areas and periods with the highest weighted scores, potentially reducing the frequency and severity of shootings by aligning law enforcement presence with predicted incident peaks.

@@

This study offers a detailed examination of the spatial and temporal factors influencing the severity of shootings and firearm discharges in Toronto, employing a robust Gradient Boosting Machine model. The findings highlight the critical role of specific neighborhoods and time periods in the incidence of these violent events, underscoring the potential for targeted interventions. By focusing law enforcement efforts and community resources on high-risk areas and times identified through the analysis, it is possible to significantly mitigate the impact of such incidents on public safety. The actionable insights provided by this study not only serve as a guide for strategic policy formulation and resource allocation but also support the ongoing efforts to enhance the efficacy of public safety measures in urban environments. This approach ensures that policy responses are data-driven and tailored to the unique dynamics of the city, ultimately contributing to a safer community for all residents.
